---
title: "Lab 10: Filtering Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you will again have a choice of three datasets to work with. You can
choose either to investigate the daily team ELO ratings in the NBA:

```{r}
nba <- read_csv("https://statsmaths.github.io/stat_data/nba_elo_daily.csv")
```











## Instructions

This lab is similar to the last. You'll pick a dataset and then generate an
analysis of that data using data visualizations. The difference here is that
instead of a single plot, I would like you to construct 3-4 plots that
together tell a linear story. Each plot should be separated by a sentence or
two describing what the viewer should take away from the plot. Try to keep the
plots similar in some way; perhaps each simply highlights a different subset
of the data but has the same underlying layers.

Notice that each of the datasets for today are larger than you will probably
be able to use. You will need to filter the data to a particular team, year,
subset of curse words. You may further filter the data to highlight an even
smaller subset of the data within each plot.

## Analysis


```{r}
new_nba <- filter(nba, team == "BRK")
new_nba

```

```{r}
ggplot(new_nba, aes(season, elo)) +
  geom_point()
```
This plot takes a closer look at the history of the elo of the Brooklyn Nets.

```{r}
best <- filter(nba, elo > 1499)
best
ggplot(new_nba, aes(season, elo)) +
  geom_point() +
  geom_point(color = "red", data = best)

```
This graph shows the times the Nets(black) dipped below the nba average elo.  

```{r}
worst <- filter(nba, elo < 1500)
best
ggplot(new_nba, aes(season, elo)) +
  geom_point() +
  geom_point(color = "pink", data = worst)
```
This graph shows all the times the Nets were above the league avg elo.
```{r}
ggplot(new_nba, aes(season, elo))+ 
  geom_point()+
  annotate("text", x = 2013, y = 1587.034, label = "Best Elo") +
  annotate("point",x = 2013, y = 1587.034, color = "red") 

```
This plot shows the greatest elo experienced in the Nets breif history (1587.034).  The point is highlighted in red and labeled.  This season player Williams and Johnson were in their prime. 


  


